---
title: 'ML #3 Predictive Policing'
author: "Prof. Fichman & Prof. Harris"
date: "08/15/2022"
output: html_document
---

We are going to run through the code base with just a couple variables in a model - in a slightly simplified workflow

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)

library(tidyverse)
library(sf)
library(RSocrata)
library(viridis)
library(spatstat)
library(raster)
library(spdep)
library(FNN)
library(grid)
library(gridExtra)
library(knitr)
library(kableExtra)
library(tidycensus)
library(classInt)   # for KDE and ML risk class intervals
# functions
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

```

## Read in Data from Chicago

This uses the Socrata package for some data sets.

Note where we bring in burglary data - you will want to vary this part to do your homework!

```{r}
policeDistricts <- 
  st_read("https://data.cityofchicago.org/api/geospatial/fthy-xz3r?method=export&format=GeoJSON") %>%
  st_transform('ESRI:102271') %>%
  dplyr::select(District = dist_num)
  
policeBeats <- 
  st_read("https://data.cityofchicago.org/api/geospatial/aerh-rz74?method=export&format=GeoJSON") %>%
  st_transform('ESRI:102271') %>%
  dplyr::select(District = beat_num)

bothPoliceUnits <- rbind(mutate(policeDistricts, Legend = "Police Districts"), 
                         mutate(policeBeats, Legend = "Police Beats"))

burglaries <- 
  read.socrata("https://data.cityofchicago.org/Public-Safety/Crimes-2017/d62x-nvdr") %>% 
    filter(Primary.Type == "BURGLARY" & Description == "FORCIBLE ENTRY") %>%
    mutate(x = gsub("[()]", "", Location)) %>%
    separate(x,into= c("Y","X"), sep=",") %>%
    mutate(X = as.numeric(X),Y = as.numeric(Y)) %>% 
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant")%>%
    st_transform('ESRI:102271') %>% 
    distinct()

chicagoBoundary <- 
  st_read(file.path(root.dir,"/Chapter5/chicagoBoundary.geojson")) %>%
  st_transform('ESRI:102271') 
```

## visualizing point data

Plotting point data and density

> How do we analyze point data?

> Are there other geometries useful to represent point locations?

```{r fig.width=6, fig.height=4}
# uses grid.arrange to organize independent plots
grid.arrange(ncol=2,
ggplot() + 
  geom_sf(data = chicagoBoundary) +
  geom_sf(data = burglaries, colour="red", size=0.1, show.legend = "point") +
  labs(title= "Burlaries, Chicago - 2017") +
  mapTheme(title_size = 14),

ggplot() + 
  geom_sf(data = chicagoBoundary, fill = "grey40") +
  stat_density2d(data = data.frame(st_coordinates(burglaries)), 
                 aes(X, Y, fill = ..level.., alpha = ..level..),
                 size = 0.01, bins = 40, geom = 'polygon') +
  scale_fill_viridis() +
  scale_alpha(range = c(0.00, 0.35), guide = FALSE) +
  labs(title = "Density of Burglaries") +
  mapTheme(title_size = 14) + theme(legend.position = "none"))
```

## Creating a fishnet grid

> What is a fishnet grid?

The `{sf}` package offers really easy way to create fishnet grids using the `st_make_grid()` function. The `cellsize` argument allows you to set the size of the grid cells; in this case it is set to `500` meters. You may have to do some research on the spatial layers projection (using `st_crs()` to know what coordinate system you are in) to understand if you are in feet or meters. If you are using Longitude and Latitude, you will need to project the data to a projected coordinate system to get distance measurements.

Examine the fishnet - the unique ID is crucial to building a data set!

```{r}
## using {sf} to create the grid
## Note the `.[chicagoBoundary] %>% ` line. This is needed to clip the grid to our data
fishnet <- 
  st_make_grid(chicagoBoundary,
               cellsize = 500, 
               square = TRUE) %>%
  .[chicagoBoundary] %>%            # fast way to select intersecting polygons
  st_sf() %>%
  mutate(uniqueID = 1:n())


```

### Aggregate points to the fishnet

> How can we aggregate points into a fishnet grid?

```{r}
## add a value of 1 to each crime, sum them with aggregate
crime_net <- 
  dplyr::select(burglaries) %>% 
  mutate(countBurglaries = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countBurglaries = replace_na(countBurglaries, 0),
         uniqueID = 1:n(),
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE))

ggplot() +
  geom_sf(data = crime_net, aes(fill = countBurglaries), color = NA) +
  scale_fill_viridis() +
  labs(title = "Count of Burglaires for the fishnet") +
  mapTheme()

# For demo. requires updated mapview package
# xx <- mapview::mapview(crime_net, zcol = "countBurglaries")
# yy <- mapview::mapview(mutate(burglaries, ID = seq(1:n())))
# xx + yy
```

## Modeling Spatial Features

> What features would be helpful in predicting the location of burglaries?
>
> What might these features be problematic?
>
> hint: for all the reasons we learned in class

```{r}
## only pulling a single variable for our model to keep it simple
## using Socrata again
abandonCars <- 
  read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Abandoned-Vehicles/3c9v-pnva") %>%
    mutate(year = substr(creation_date,1,4)) %>% filter(year == "2017") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Abandoned_Cars")

## Neighborhoods to use in LOOCV in a bit
neighborhoods <- 
  st_read("https://raw.githubusercontent.com/blackmad/neighborhoods/master/chicago.geojson") %>%
  st_transform(st_crs(fishnet)) 

```

#### How we aggregate a feature to our fishnet

This is an important chunk of code with some unfamiliar lines. The overall to assign the fishnet ID to each abandoned car point, group the points by fishnet ID and count them per ID, join that count back to the fishnet and then go from a long format to a wide format. We'll step through it:

*   `st_join(fishnet, join=st_within)`
*       spatially join `abandonCars` points to the `fishnet` polygon they are within by specifying `st_within`. This results in the `Abandoned_Cars` points being given the `uniqueID` of the fishnet grid cell that they are within.
*   `st_drop_geometry()`
*       drop the geometry attributes of the joined data so that we can do summaries on it without having to also calculate geometries which would be quite slow.
*   `group_by(uniqueID, Legend)`
*       we want to count the number of abandoned cars per fishnet, so we use `group_by` on the unique cell ID. We also include the `Legend` column, which is more useful if you are doing this one more than a single layer.
*   `summarize(count = n())`
*       use `summarize` to create a new field called `count` that will be the count of all abandoned cars per fishnet grid cell. The `n()` function returns the number within each group (i.e. `uniqueID`)
*   `left_join(fishnet, ., by = "uniqueID")`
*       join that summary back to spatial fishnet by the `uniqueID` field which is in both. Note the use of the `.` "dot operator" which is a stand in for the object being piped into the function (i.e. `left_join()`). We use this because we want the summaries to be the second argument of the join; not the first argument which is the dplyr default.
*   `spread(Legend, count, fill=0)`
*       "spread" from long to wide format and make a new column for each value of the `Legend` field. This also is more useful if there are multiple layers in the `Legend` column. Note the use of `fill=0` tells the function to fill in any fishnet cells without an abandoned car with a `0` instead of `NA`.
*   `dplyr::select(-``<NA>``)`
*       remove a `<NA>` column that was created because of the `NA` value in the `Legend` column when it was "spread"
*   `ungroup()`
*       Finally, ungroup the dataframe.

```{r}

vars_net <- abandonCars %>%
  st_join(fishnet, join=st_within) %>%
  st_drop_geometry() %>%
  group_by(uniqueID, Legend) %>%
  summarize(count = n()) %>%
  left_join(fishnet, ., by = "uniqueID") %>%
  spread(Legend, count, fill=0) %>%
  dplyr::select(-`<NA>`) %>%
  ungroup()


```

## Nearest Neighbor Feature

> Review: what is NN and what does `k` represent in this function?

```{r}
# convinience to reduce length of function names.
st_c    <- st_coordinates
st_coid <- st_centroid

## create NN from abandoned cars
vars_net <- vars_net %>%
    mutate(Abandoned_Cars.nn = nn_function(st_c(st_coid(vars_net)), 
                                           st_c(abandonCars),
                                           k = 3))
```

> What changes if we make `k` a different number?

```{r}
## Visualize the NN feature
vars_net.long.nn <- 
  dplyr::select(vars_net, ends_with(".nn")) %>%
    gather(Variable, value, -geometry)

ggplot() +
      geom_sf(data = vars_net.long.nn, aes(fill=value), colour=NA) +
      scale_fill_viridis(name="NN Distance") +
      labs(title="Abandoned Car NN Distance") +
      mapTheme()
```

## Join NN feature to our fishnet

Since the counts were aggregated to each cell by `uniqueID` we can use that to join the counts to the fishnet.

```{r}
## important to drop the geometry from joining features
final_net <-
  left_join(crime_net, st_drop_geometry(vars_net), by="uniqueID") 

```

### Join in areal data

Using spatial joins to join *centroids* of fishnets to polygon for neighborhoods and districts.

> What issues arise when we try to join polygons to polygons in space?

```{r}

final_net <-
  st_centroid(final_net) %>%
    st_join(dplyr::select(neighborhoods, name), by = "uniqueID") %>%
    st_join(dplyr::select(policeDistricts, District), by = "uniqueID") %>%
      st_drop_geometry() %>%
      left_join(dplyr::select(final_net, geometry, uniqueID)) %>%
      st_sf() %>%
  na.omit()

# for live demo
# mapview::mapview(final_net, zcol = "District")
```

## Local Moran's I for fishnet grid cells

using {spdep} package to to build neighborhood weights and list to calculate local Moran's I.

Note that the code here is *different* than in the book - it has been updated to keep up with changes in packages.

> What is the difference between local and global Moran's I?

A little in depth version of the chunk below can be found:

Mendez C. (2020). Spatial autocorrelation analysis in R. R Studio/RPubs. Available at <https://rpubs.com/quarcs-lab/spatial-autocorrelation>

```{r}
## generates warnings from PROJ issues
## {spdep} to make polygon to neighborhoods... 
final_net.nb <- poly2nb(as_Spatial(final_net), queen=TRUE)
## ... and neighborhoods to list of weigths
final_net.weights <- nb2listw(final_net.nb, style="W", zero.policy=TRUE)

# print(final_net.weights, zero.policy=TRUE)
```

```{r}
## see ?localmoran
local_morans <- localmoran(final_net$Abandoned_Cars, final_net.weights, zero.policy=TRUE) %>% 
  as.data.frame()

# join local Moran's I results to fishnet
final_net.localMorans <- 
  cbind(local_morans, as.data.frame(final_net)) %>% 
  st_sf() %>%
  dplyr::select(Abandoned_Cars_Count = Abandoned_Cars, 
                Local_Morans_I = Ii, 
                P_Value = `Pr(z != E(Ii))`) %>%
  mutate(Significant_Hotspots = ifelse(P_Value <= 0.001, 1, 0)) %>%
  gather(Variable, Value, -geometry)
  
```

### Plotting local Moran's I results

This is a complex code chunk - it's a loop which builds ggplots of local Moran's for each of your `vars`

> What does a significant hot spot tell us about the distribution of burglaries?

```{r}
## This is just for plotting
vars <- unique(final_net.localMorans$Variable)
varList <- list()

for(i in vars){
  varList[[i]] <- 
    ggplot() +
      geom_sf(data = filter(final_net.localMorans, Variable == i), 
              aes(fill = Value), colour=NA) +
      scale_fill_viridis(name="") +
      labs(title=i) +
      mapTheme(title_size = 14) + theme(legend.position="bottom")}

do.call(grid.arrange,c(varList, ncol = 4, top = "Local Morans I statistics, Burglary"))
```

## Distance to Hot spot

Using NN distance to a hot spot location

```{r}
# generates warning from NN
final_net <- final_net %>% 
  mutate(abandoned.isSig = 
           ifelse(local_morans[,5] <= 0.001, 1, 0)) %>%
  mutate(abandoned.isSig.dist = 
           nn_function(st_c(st_coid(final_net)),
                       st_c(st_coid(filter(final_net, 
                                           abandoned.isSig == 1))), 
                       k = 1))

## What does k = 1 represent?
```

> What does `k = 1` above mean in terms of measuring nearest neighbors?

### Plot NN distance to hot spot

```{r}
ggplot() +
      geom_sf(data = final_net, aes(fill=abandoned.isSig.dist), colour=NA) +
      scale_fill_viridis(name="NN Distance") +
      labs(title="Abandoned Car NN Distance") +
      mapTheme()
```

## Modeling and CV

Leave One Group Out CV on spatial features

```{r sults='hide'}

# View(crossValidate)

## define the variables we want
reg.ss.vars <- c("Abandoned_Cars.nn", "abandoned.isSig.dist")

## RUN REGRESSIONS
reg.ss.spatialCV <- crossValidate(
  dataset = final_net,
  id = "name",                           
  dependentVariable = "countBurglaries",
  indVariables = reg.ss.vars) %>%
    dplyr::select(cvID = name, countBurglaries, Prediction, geometry)
```

```{r}
# calculate errors by NEIGHBORHOOD
error_by_reg_and_fold <- 
  reg.ss.spatialCV %>%
    group_by(cvID) %>% 
    summarize(Mean_Error = mean(Prediction - countBurglaries, na.rm = T),
              MAE = mean(abs(Mean_Error), na.rm = T),
              SD_MAE = mean(abs(Mean_Error), na.rm = T)) %>%
  ungroup()

error_by_reg_and_fold %>% 
  arrange(desc(MAE))
error_by_reg_and_fold %>% 
  arrange(MAE)

## plot histogram of OOF (out of fold) errors
error_by_reg_and_fold %>%
  ggplot(aes(MAE)) + 
    geom_histogram(bins = 30, colour="black", fill = "#FDE725FF") +
  scale_x_continuous(breaks = seq(0, 11, by = 1)) + 
    labs(title="Distribution of MAE", subtitle = "LOGO-CV",
         x="Mean Absolute Error", y="Count") 
```

## Density vs predictions

The `spatstat` function gets us kernal density estimates with varying search radii.

Note that the code here is *different* than in the book - it has been updated to keep up with changes in packages.

```{r}
# demo of kernel width
burg_ppp <- as.ppp(st_coordinates(burglaries), W = st_bbox(final_net))
burg_KD.1000 <- spatstat.core::density.ppp(burg_ppp, 1000)
burg_KD.1500 <- spatstat.core::density.ppp(burg_ppp, 1500)
burg_KD.2000 <- spatstat.core::density.ppp(burg_ppp, 2000)
burg_KD.df <- rbind(
  mutate(data.frame(rasterToPoints(mask(raster(burg_KD.1000), as(neighborhoods, 'Spatial')))), Legend = "1000 Ft."),
  mutate(data.frame(rasterToPoints(mask(raster(burg_KD.1500), as(neighborhoods, 'Spatial')))), Legend = "1500 Ft."),
  mutate(data.frame(rasterToPoints(mask(raster(burg_KD.2000), as(neighborhoods, 'Spatial')))), Legend = "2000 Ft.")) 

burg_KD.df$Legend <- factor(burg_KD.df$Legend, levels = c("1000 Ft.", "1500 Ft.", "2000 Ft."))

ggplot(data=burg_KD.df, aes(x=x, y=y)) +
  geom_raster(aes(fill=layer)) + 
  facet_wrap(~Legend) +
  coord_sf(crs=st_crs(final_net)) + 
  scale_fill_viridis(name="Density") +
  labs(title = "Kernel density with 3 different search radii") +
  mapTheme(title_size = 14)
```

```{r}

as.data.frame(burg_KD.1000) %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(final_net)) %>%
  aggregate(., final_net, mean) %>%
   ggplot() +
     geom_sf(aes(fill=value)) +
     geom_sf(data = sample_n(burglaries, 1500), size = .5) +
     scale_fill_viridis(name = "Density") +
     labs(title = "Kernel density of 2017 burglaries") +
     mapTheme(title_size = 14)
```

## Get 2018 crime data

Let's see how our model performed relative to KD on the following year's data.

```{r}
burglaries18 <- 
  read.socrata("https://data.cityofchicago.org/Public-Safety/Crimes-2018/3i3m-jwuy") %>% 
  filter(Primary.Type == "BURGLARY" & 
         Description == "FORCIBLE ENTRY") %>%
  mutate(x = gsub("[()]", "", Location)) %>%
  separate(x,into= c("Y","X"), sep=",") %>%
  mutate(X = as.numeric(X),
         Y = as.numeric(Y)) %>% 
  na.omit %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform('ESRI:102271') %>% 
  distinct() %>%
  .[fishnet,]
```

```{r}

burg_KDE_sum <- as.data.frame(burg_KD.1000) %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(final_net)) %>%
  aggregate(., final_net, mean) 
kde_breaks <- classIntervals(burg_KDE_sum$value, 
                             n = 5, "fisher")
burg_KDE_sf <- burg_KDE_sum %>%
  mutate(label = "Kernel Density",
         Risk_Category = classInt::findCols(kde_breaks),
         Risk_Category = case_when(
           Risk_Category == 5 ~ "5th",
           Risk_Category == 4 ~ "4th",
           Risk_Category == 3 ~ "3rd",
           Risk_Category == 2 ~ "2nd",
           Risk_Category == 1 ~ "1st")) %>%
  cbind(
    aggregate(
      dplyr::select(burglaries18) %>% mutate(burgCount = 1), ., sum) %>%
    mutate(burgCount = replace_na(burgCount, 0))) %>%
  dplyr::select(label, Risk_Category, burgCount)
```

Note that this is different from the book, where we pull a model out of a list of models we've created. For your homework, you'll be creating multiple models.

```{r}
ml_breaks <- classIntervals(reg.ss.spatialCV$Prediction, 
                             n = 5, "fisher")
burg_risk_sf <-
  reg.ss.spatialCV %>%
  mutate(label = "Risk Predictions",
         Risk_Category =classInt::findCols(ml_breaks),
         Risk_Category = case_when(
           Risk_Category == 5 ~ "5th",
           Risk_Category == 4 ~ "4th",
           Risk_Category == 3 ~ "3rd",
           Risk_Category == 2 ~ "2nd",
           Risk_Category == 1 ~ "1st")) %>%
  cbind(
    aggregate(
      dplyr::select(burglaries18) %>% mutate(burgCount = 1), ., sum) %>%
      mutate(burgCount = replace_na(burgCount, 0))) %>%
  dplyr::select(label,Risk_Category, burgCount)
```

We don't do quite as well because we don't have very many features, but still pretty good.

```{r}
rbind(burg_KDE_sf, burg_risk_sf) %>%
  na.omit() %>%
  gather(Variable, Value, -label, -Risk_Category, -geometry) %>%
  ggplot() +
    geom_sf(aes(fill = Risk_Category), colour = NA) +
    geom_sf(data = sample_n(burglaries18, 3000), size = .5, colour = "black") +
    facet_wrap(~label, ) +
    scale_fill_viridis(discrete = TRUE) +
    labs(title="Comparison of Kernel Density and Risk Predictions",
         subtitle="2017 burglar risk predictions; 2018 burglaries") +
    mapTheme(title_size = 14)
```

```{r}
rbind(burg_KDE_sf, burg_risk_sf) %>%
  st_drop_geometry() %>%
  na.omit() %>%
  gather(Variable, Value, -label, -Risk_Category) %>%
  group_by(label, Risk_Category) %>%
  summarize(countBurglaries = sum(Value)) %>%
  ungroup() %>%
  group_by(label) %>%
  mutate(Pcnt_of_test_set_crimes = countBurglaries / sum(countBurglaries)) %>%
    ggplot(aes(Risk_Category,Pcnt_of_test_set_crimes)) +
      geom_bar(aes(fill=label), position="dodge", stat="identity") +
      scale_fill_viridis(discrete = TRUE, name = "Model") +
      labs(title = "Risk prediction vs. Kernel density, 2018 burglaries",
           y = "% of Test Set Burglaries (per model)",
           x = "Risk Category") +
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```
