---
title: "Lab 1 - Learning Tidy R With Census Data"
author: "Michael Fichman"
date: "MUSA 508, Fall, 2021"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(kableExtra)
library(rmarkdown)
```

# Introduction

This tutorial will introduce you to the use of R for US Census data. This code uses the `tidycensus` pacakge to query the US Census Bureau's Application Programming Interface (API). An API is, roughly, a "place" on the internet which you can interact with using a computer program. In our case, we will interact with an API which allows us access to US Census Data - this is the same data library that the Census' own website uses.

Perhaps more importantly, this tutorial introduces you to the `tidyverse` world of libraries - the best data wrangling, summarizing and visualizing tools available in R. These packages are designed to "play nice" with each other - none of them use the same function names and they are maintained together by a team of developers to mkae sure they work well.

R Studio publishes handy "cheat sheets" for popular R libraries, including [data wrangling in the tidyverse](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf). 

This exercise consists of a few classic planner's tasks:

1. Collecting and manipulating census data

2. Measuring demographic change over time for a given geography

3. Measuring statistics for a geography relative to a parent geography

4. Creating summary statistics regarding demographics for a collection of geographies

# Setup 

## Install Libraries

If you haven't installed either `tidyverse`, `tidycensus`, or `sf` use the `install.packages` command like so:

(If you have them installed, you can skip this step)

```{r setup_packages1, warning = FALSE, eval = FALSE}
install.packages('tidyverse')
install.packages('tidycensus')
install.packages('sf')
```

Once the packages are installed, you must load them using hte `library` command so that they are active in your environment.

```{r setup_packages2, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidycensus)
library(sf)
```

## Census API Key

You will need a "key" to access the Census API. You can find one at [their website](https://api.census.gov/data/key_signup.html).

Paste it into the code block below:

```{r load_key_hide, warning= FALSE, include=FALSE}
census_api_key("e79f3706b6d61249968c6ce88794f6f556e5bf3d", overwrite = TRUE)
```

```{r load_key, warning = FALSE, eval = FALSE}
census_api_key("YOUR KEY GOES HERE", overwrite = TRUE)
```

## Load census data dictionaries

Now that we have our census credentials loaded, we can start downloading information from the API using some functions from tidycensus. We are going to make some comparisons between 2016 and 2010 ACS estimates for our target census tracts. In order to choose variables of interest, we are going to load the data dictionaries for each time period using `load_variables`.

What does this `load_variables` function do and why are we writing these things in the parentheses after the name of the function? If you want to know about a function in R, you can type the name of a package or function into the console like this: `??load_variables` and information will show up in the help window in your R Studio environment.

```{r load_variables, cache = TRUE}
acs_variable_list.2016 <- load_variables(2016, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)

acs_variable_list.2010 <- load_variables(2010, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)
```

Once we have loaded these data frames, we can observe and search through the data frames of variable information which should appear in our global environment either by clicking on them or using the `View(YOUR DATAFRAME NAME HERE)` command.

Let's look around in these data frames for a few minutes and see what's in there.

# Downloading Data from Tidycensus

## Create a vector of census variables

We can populate a vector of variable names we will send to the Census API. We call this list `acs_vars`. This is the beauty of a code-based workflow - you can take this vector and put anything you want in it when you have a new analysis to do and re-run it for different variables. These need to be character strings, and hence, in quotes as you see below.

Keep in mind the categories and code numbers change a bit over time - you may need separate vectors for different census years.

```{r acs_vars}
acs_vars <- c("B01001_001E", # ACS total Pop estimate
              "B25002_001E", # Estimate of total housing units
              "B25002_003E", # Number of vacant housing units
              "B19013_001E", # Median HH Income ($)
              "B02001_002E", # People describing themselves as "white alone"
              "B06009_006E") # Total graduate or professional degree
```

## Call the Census API to get tract level data for 2016 for all of Philadelphia

We use the `get_acs` function in `tidycensus` to query the API. Notice the different arguments for the function, and that they require certain types of info. For example, `geography` requires one of a finite list of answers, and they have to be formatted as character string.

Remember the `??` function - you can learn about the parameters for `get_acs` this way. There is also a function called `get_decennial` which you can use for decennial census counts.

We ask for data on our `acs_vars` for all tracts in Philadelphia County, PA in 2016. We ask for "wide" data (e.g. one variable per column, one row per tract) and we set `geometry` to `FALSE`.

What happens when you set `geometry` to `TRUE`?

```{r get_acs_2016, cache = TRUE, message = FALSE, warning = FALSE}
acsTractsPHL.2016 <- get_acs(geography = "tract",
                             year = 2016, 
                             variables = acs_vars, 
                             geometry = FALSE, 
                             state = "PA", 
                             county = "Philadelphia", 
                             output = "wide") 
```

# Wrangling Data with dplyr

## Mutating, selecting and renaming variables

Now we can manipulate our data using some of our `tidyverse` data wrangling tools from the `dplyr` library.

The `dplyr` package is great for these operations and has some very common sense functions that are fairly intuitive because they use "verbs". You can -

-select columns (`select`)

-rename columns (`rename`)

-summarize data (`summarize`) by groups (`group_by`)

- create new columns and specify their value (`mutate`)

The operator `%>%` is known as the "pipe" and lets you chain operations together - passing a dataframe along through different operations.

Let's manipulate our data using the "pipe" %>% and some tidy data wrangling commands. We will remove some variables using the "select" command and a parenthesis preceded by a minus sign. We will rename some variables using "rename". We will create some new ones using "mutate"

First we select tract `GEOID` and `NAME` and only those variables in the `acs_vars` list - we are only using the estimates from the ACS (suffix "E" on the variable names), we are not retaining the margin of error calculations (suffix "M" on the variable names).

Look at the output and see what has changed.

Second we `rename` our variables to more common sense names.

Lastly we `mutate` a few new columns.

```{r do_some_dplyr, cache = TRUE}
acsTractsPHL.2016 <- acsTractsPHL.2016 %>%
  dplyr::select (GEOID, NAME, all_of(acs_vars))

acsTractsPHL.2016 <- acsTractsPHL.2016 %>%
  rename (total_pop.2016 = B01001_001E,
          total_HU.2016 = B25002_001E,
          total_vacant.2016 = B25002_003E,
          med_HH_Income.2016 = B19013_001E,
          total_White.2016 = B02001_002E,
          total_GradDeg.2016 = B06009_006E)

acsTractsPHL.2016 <- acsTractsPHL.2016 %>%
  mutate(vacancyPct.2016 = total_vacant.2016/total_HU.2016,
         pctWhite.2016 = total_White.2016/total_pop.2016)
```

Let's grab the 2010 data and do the same operations, but now - let's *chain the piped operations directly to the API call and do it all at once*. The only difference here is that we are going to change the variable names to say "2010" in them, not "2016".

```{r get_acs_2010, cache = TRUE, message = FALSE}
acsTractsPHL.2010 <- get_acs(geography = "tract",
                             year = 2010, 
                             variables = acs_vars,
                             geometry = FALSE,
                             state = "PA", 
                             county = "Philadelphia",
                             output = "wide") %>%
  dplyr::select (GEOID, NAME, all_of(acs_vars)) %>% 
  rename (total_pop.2010 = B01001_001E,
          total_HU.2010 = B25002_001E,
          total_vacant.2010 = B25002_003E,
          med_HH_Income.2010 = B19013_001E,
          total_White.2010 = B02001_002E,
          total_GradDeg.2010 = B06009_006E) %>%
  mutate(vacancyPct.2010 = total_vacant.2010/total_HU.2010,
         pctWhite.2010 = total_White.2010/total_pop.2010)
```

## Joining data

We can do a tabular join - each of our census dataframes should have a unique ID for each tract. You need to join columns that consist of the same type of data. Using hte `glimpse` command can give you a quick overview of the data types for each column in your data set. Recall that joining character to numeric data is going to be a no-go.

There are four kinds of joins - you can learn about them here: https://rpubs.com/williamsurles/293454

We are going to do a "left join," keeping all the observations from the left hand side of the join argument, and only those which match from the right. Ideally this consists of all of the observations from the right hand side considering we have the same geometries in 2010 and 2016.

```{r left_join_tracts, cache = TRUE}
allACS <- left_join(acsTractsPHL.2010, acsTractsPHL.2016, by= c("GEOID"))
```

## Doing column math using mutate

The `mutate` function can be used to create new columns. These columns can be the product of arithmetic using other columns. Here we are going to calculate changes in inflation-adjusted median household income from 2010-2016 and the change in the percentage of people holding a graduate degree. (Notice how we multiply `med_HH_Income.2010` by 1.10007 to adjust it to 2016 dollars).

Notice that you are overwriting allACS with a new data frame called the same thing. You can overwrite anything if you create something with the same name - so be careful!

```{r do_mutates, cache = TRUE}
allACS <- allACS %>%
  mutate(change_med_HH_Income = med_HH_Income.2016 - (med_HH_Income.2010 * 1.1007), 
         change_Grad_Degree_Pct = (total_GradDeg.2016/total_pop.2016)-(total_GradDeg.2010/total_pop.2010))

```

## Exercise - Creating New Variables

Use the `mutate` command to create three new variables in your allACS data set. These variables can either be measures of change or proportions/percentages.

If you want, you can put some new variables in your ACS data sets by adding to `acs_vars` - just make sure you put something in the `rename` commands we used earlier if you want to have intelligible variable names.

# Summarizing Census Data

## Exploring central tendancies

Base R language has lots of summary statistical functions, like `mean` and `median`. We can apply these to columns in our data frame.

This one comes back `NA`. NA plus NA equals NA - so summarizing these data won't work.

```{r base_r_summaries_NA_example, cache = TRUE}
mean(allACS$change_med_HH_Income)
```

OK, let's try it again, using a workaround. Can you think of some reasons why omitting NA data might be problematic?

```{r base_r_summaries_NA_rm, cache = TRUE}
mean(allACS$change_med_HH_Income, na.rm = TRUE)

median(allACS$change_med_HH_Income, na.rm = TRUE)
```

## Exploring distributions

Base R has a graphics package - we can look at the distribution of the data, instead of just the central tendencies (e.g. `mean` and `median`).

```{r basic_histogram, cache = TRUE}
hist(allACS$change_med_HH_Income)
```

We can also use the dynamic graphics package from the tidyverse, `gglot2` to make some more detailed histograms. We have lots of aesthetic options here.

Make your own by changing the terms in the ggplot call - use a different variable and change the axis titles. You see we can build a nice plot with different elements using building blocks like `geom_histogram`.

Notice that the `+` sign does the same work in ggplot that the `%>%` does in tidy code.

```{r ggplot_histogram_simple, warning = FALSE, cache = TRUE, message = FALSE}
ggplot(allACS)+
  geom_histogram(aes(change_med_HH_Income))
  
```

Let's make this a bit more complex, and add some labels.

```{r ggplot_histogram, warning = FALSE, cache = TRUE}
ggplot(allACS)+
  geom_histogram(aes(change_med_HH_Income), binwidth = 5000)+
  labs(
    title = "Change in Philadelphia HH median income by tract, 2010-2016",
    caption = "Data: US Census Bureau, ACS 5-year estimates",
    x="Change in Med HH Income (2016 dollars)", 
       y="Number of tracts")
  
```

Wow, there are some tracts with a huge increase in Med HH Income! What is the simplest explanation for that?

## Making a summary table

The `summarize` function from the `dplyr` library is very powerful - you can use these summary functions in the context of a table - this is akin to "pivot tables" and other tools from Excel - but this is incredibly simple. You could create a table of statistics and easily export it for use in a report.

```{r summaryTable, cache = TRUE}
summaryTable <- allACS %>%
  summarize(mean_change_HH_Income = mean(change_med_HH_Income, na.rm = TRUE),
            med_change_HH_Income = median(change_med_HH_Income, na.rm = TRUE))
```


# Comparing geographies

Let's say you are a planner working for the City, and you are creating a neighborhood plan for a particular neighborhood. Our example neighborhood will be Mt. Airy, in Philadelphia's Northwest section.

Let's start by making a vector that has the GEOIDs of census tracts in Mt. Airy. The allACS$GEOID column is a CHARACTER, not a number, athough it appears numeric.

We can then create a vactor variable, which we call `mtAiry`, using a "boolean" statement.

We say that if a tract has a GEOID in `myTracts`, the variable `mtAiry` should take on a value of "MT AIRY", or else it takes on a value of "REST OF PHILADELPHIA"

```{r myTracts, cache = TRUE}

myTracts <- c("42101023500", 
              "42101023600", 
              "42101023700", 
              "42101025300", 
              "42101025400",
              "42101025500", 
              "42101025600", 
              "42101038800")

allACS <- allACS %>%
  mutate(mtAiry = ifelse(GEOID %in% myTracts, "MT AIRY", "REST OF PHILADELPHIA"))
```

Let's make a new summary table, this time accounting for the statistics for both Mt. Airy and the rest of the city. 

We can see that city-wide, "on average" the mean tract income level increased, while the median level decreased. Why might this be?

We had the opposite dynamic in Mt. Airy, where median household income per tract rose, while mean income fell.

```{r summary_table_2, cache = TRUE, message = FALSE, warning = FALSE}
summaryTable2 <- allACS %>%
  group_by(mtAiry) %>%
  summarize(mean_change_HH_Income = mean(change_med_HH_Income, na.rm = TRUE),
            med_change_HH_Income = median(change_med_HH_Income, na.rm = TRUE))
```

## Graphic comparisons Using ggplot2

A graphical approach to subdividing and summarizing the data might make this easier to figure out.

We could compare the distributions between Mt. Airy and the City by taking our histogram we created earlier and using the `facet_wrap` option to subdivide the data and create a side-by-side comparison.

What can we see here about the distributions of the data, and how the distribution and the central tendencies might tell us different stories about what's happening?

```{r ggplot_histogram_2, warning = FALSE, cache = TRUE}
ggplot(allACS)+
  geom_histogram(aes(change_med_HH_Income), binwidth = 10000)+
  labs(
    title = "Change in Philadelphia HH median income by tract, 2010-2016",
    caption = "Data: US Census Bureau, ACS 5-year estimates",
    x="Change in Med HH Income (2016 dollars)", 
       y="Number of tracts")+
  facet_wrap(~mtAiry, scales = "free")
  
```

We can create a scatterplot to see what the relationship is between 2010 and 2016 income and where our Mt. Airy tracts fall.

Notice how we put the inflation adjustment right in the `geom_point` call - you can temporarily create new data inside your ggplot diagrams!

I set a `fill` parameter to differentiate between our factor levels in the data.

I add a line that represents y=x using `geom_abline`.

What do you think this line represents? How does it help us interpret these points we are seeing?

```{r ggplot_point1, warning = FALSE, cache = TRUE}
ggplot(allACS)+
  geom_point(aes(x =med_HH_Income.2010 * 1.1007, 
                 y = med_HH_Income.2016,
                 color = mtAiry))+
  geom_abline(intercept = 0, slope = 1)+
  labs(
    title = "2016 Median HH Income as a Function of 2010 Median HH Income",
    subtitle = "All figures in 2016 dollars",
    caption = "Data: US Census Bureau, ACS 5-year estimates",
    x="Med HH Income 2010 ($)", 
    y="Med HH Income 2016 ($)")
  
```

There are very well established correlations between socio-demographic characteristics and wealth in the US. Philadelphia is no exception. A simple scatterplot is stark evidence of this correlation. 

We can plot a `geom_smooth` trendline in here using the `method = "lm"` e.g. a linear fit to see.

How do our tracts in Mt. Airy look in the context of distribution of wealth and relative to the trend line?

If you plot these data for 2010, do they look different?


```{r ggplot_point2, warning = FALSE, cache = TRUE, message = FALSE}
ggplot(allACS)+
  geom_point(aes(x = 100* pctWhite.2016, 
                 y = med_HH_Income.2016,
                 color = mtAiry))+
  geom_smooth(aes(x = 100* pctWhite.2016, 
                  y = med_HH_Income.2016), 
              method = "lm", se = FALSE)+
  labs(
    title = "2016 Median HH Income as a Function of Pct White",
    subtitle = "All figures in 2016 dollars",
    caption = "Data: US Census Bureau, ACS 5-year estimates",
    x="Pct. Residents Identifying as 'White Only'", 
    y="Med HH Income 2016 ($)")
  
```

# Spatial Data and Tidycensus

Tidycensus can also provide us with a geometry for our data. The `sf` package handles shapefiles and interfaces well with `tidycensus` to allow us to analyze and map spatial data. Next lesson we will do some spatial analysis using these data.

Let's re-run our call for 2016 Census Data and turn the `geometry = TRUE` and then we can make a map of `pctWhite.2016` using the tidycensus data right in ggplot using a special `geom`.

Notice that there is a `geometry` column in our data frame now - it's actually an `sf` object, which is just like a data frame except it has a column with drawing instructions that you can't operate on or alter.

Observe the use of the `%>%` operator here - I'm going to chain all of our operations together into one big chunk of code:

```{r spatial_tidycensus, message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
acsTractsPHL.2016.sf <- get_acs(geography = "tract",
                             year = 2016, 
                             variables = acs_vars, 
                             geometry = TRUE, 
                             state = "PA", 
                             county = "Philadelphia", 
                             output = "wide") %>% 
  dplyr::select (GEOID, NAME, all_of(acs_vars)) %>%
  rename (total_pop.2016 = B01001_001E,
          total_HU.2016 = B25002_001E,
          total_vacant.2016 = B25002_003E,
          med_HH_Income.2016 = B19013_001E,
          total_White.2016 = B02001_002E,
          total_GradDeg.2016 = B06009_006E) %>%
  mutate(vacancyPct.2016 = total_vacant.2016/total_HU.2016,
         pctWhite.2016 = total_White.2016/total_pop.2016) %>%
  mutate(mtAiry = ifelse(GEOID %in% myTracts, "MT AIRY", "REST OF PHILADELPHIA")) %>%
  st_as_sf(crs = 4326) # Turn shp into sf object and project as WGS84
```

```{r spatial_tidycensus_no_eval, message=FALSE, warning=FALSE, cache=TRUE, eval = FALSE}
acsTractsPHL.2016.sf <- get_acs(geography = "tract",
                             year = 2016, 
                             variables = acs_vars, 
                             geometry = TRUE, 
                             state = "PA", 
                             county = "Philadelphia", 
                             output = "wide") %>% 
  dplyr::select (GEOID, NAME, all_of(acs_vars)) %>%
  rename (total_pop.2016 = B01001_001E,
          total_HU.2016 = B25002_001E,
          total_vacant.2016 = B25002_003E,
          med_HH_Income.2016 = B19013_001E,
          total_White.2016 = B02001_002E,
          total_GradDeg.2016 = B06009_006E) %>%
  mutate(vacancyPct.2016 = total_vacant.2016/total_HU.2016,
         pctWhite.2016 = total_White.2016/total_pop.2016) %>%
  mutate(mtAiry = ifelse(GEOID %in% myTracts, "MT AIRY", "REST OF PHILADELPHIA")) %>%
  st_as_sf(crs = 4326) # Turn shp into sf object and project as WGS84
```

Now we can create a a `geom_sf` which will allow you to map `sf` objects in `ggplot2` so long as they are have their projection set to WGS84 (aka web mercator). You can see above that we took the shapefile created by the tidycensus call and used the `st_as_sf` call to turn it into an `sf` object and set the `crs` (coordinate reference system) to `4326`, which is the code for WGS84. You can find all these kinds of codes at [spatialreference.org](http://spatialreference.org).

Notice we create a `geom_sf` which symbologizes our data and then another which is designed to represent the boundary of Mt. Airy.

In our next lesson, we will do more detailed cartography and spatial analysis using ggplot - using custom color palettes and more.

```{r ggplot_geom_sf, warning = FALSE, cache = TRUE}
ggplot()+
  geom_sf(data = acsTractsPHL.2016.sf, aes(fill = pctWhite.2016),
          color = "transparent")+
  geom_sf(data = acsTractsPHL.2016.sf %>%
            filter(mtAiry == "MT AIRY") %>%
            st_union(),
          color = "white",
          fill = "transparent")+
  labs(
    title = "Percentage of those identifying as 'white only' by tract",
    subtitle = "",
    caption = "Data: US Census Bureau, ACS 5-year estimates")
  
```


# Discussion 

- How would you describe the income and demographic statistics and trends in Mt. Airy relative to those of Philadelphia as a whole?

- Do you have a particular hypothesis about "stories" about Mt. Airy based on a look at the data in a spatial format?

- What do you think about the similarities and differences within the neighborhood? How can you probe at those using these data?

- Can you think of any other ways to subdivide the data we looked at in class to glean more insights about Mt. Airy?